
# Overview

PowerCHORD includes three optimization methods: 
* differential evolution
* exhaustive search
* semidefinite programming
In this notebook we demonstrate how to perform an exhaustive search using PowerCHORD in R.

**Note:** We are using a very coarse grid in this notebook `(Nfine=24)` to speed up computations. For research questions, you would likely want to use a finer grid `(Nfine=48)`.
```{r}
library(PowerCHORD)
```

# Basic usage

## Example of a search

The `exhaustiveSearch` function uses an underlying `C` library to generate a database of all possible designs, and then filters for those which satisfy the window constraint using `awk`.

Schedule 8 measurements in circadian such that:
* measurements are confined to half hour grid
* there is a 12 hour window during which no measurements are scheduled

```{r}
Nfine    = 24 # corresponds to 1hr intervals in circadian study
N        = 8  # number of measurements to collect

# no measurements for half the study 
wlen     = 12  # length of window
wdensity = 0   # prescribed number of measurements during window

# if using larger Nfine, this will take a few minutes to run 
bvec = exhaustiveSearch(N=N,
                 Nfine=Nfine,wlen=wlen,wdensity=wdensity,
                 returnType='optimal')

# convert to actual measurement times
tau = c(1:Nfine)/Nfine - 1/Nfine # temporal grid
meas_times = tau[bvec>0]
evalWorstPower(meas_times,Amp=2,freq=1) 
```


## Return all constrained designs
Modify output so that all constrained designs are returned (rather than returning only the constrained-optimal design)
```{r}
Nfine    = 24 # corresponds to hour intervals in circadian study
N        = 8  # number of measurements to collect

# no measurements for half the study 
wlen     = 12  # length of window
wdensity = 0   # prescribed number of measurements during window
df = exhaustiveSearch(N=N,
                 Nfine=Nfine,wlen=wlen,wdensity=wdensity,
                 returnType='all')
nrow(df) # number of constrained solutions
head(df)
```

## Remove window constraint
If you want to see the entire database, you can turn off the window constraint as follows
```{r}
Nfine    = 24 # corresponds to hour intervals in circadian study
N        = 8  # number of measurements to collect

# no measurements for half the study 
wlen     = 1  # length of window, equivalent to removing constraint
wdensity = 1  # prescribed number of measurements during window
df = exhaustiveSearch(N=N,
                 Nfine=Nfine,wlen=wlen,wdensity=wdensity,
                 returnType='all')
nrow(df) # should be 30667 
```

Optional: if you want to count the number of designs and check that it matches
the result of exhaustive search
```{r}
#install.packages("numbers")
#require(numbers)
Nfine = 24
N     = 8
jvals = numbers::divisors(numbers::GCD(N,Nfine))
count=0
for (jj in jvals){
  count=count+numbers::eulersPhi(jj)*choose(Nfine/jj,N/jj)
}
count/Nfine
```
