```{r}
library(PowerCHORD)
```

# Power analysis

## Circadian equispaced study
Example of a circadian study with a sample size of 8 measurements. As always, measurements are scaled to lie in the interval [0,1]. 
```{r}
N    = 8              # sample size
mt   = c(1:N)/N - 1/N # vector of measurement times
mt
```

Show measurement times in circadian units (hrs)
```{r}
24*mt
```

Specify signal parameters and compare exact vs an approximate Monte Carlo power estimate
```{r}
Amp  = 2              # signal amplitude
freq = 1              # frequency
acro = pi             # acrophase
param = list(freq=freq,Amp=Amp,acro=acro)
evalExactPower(mt,param)
evalMonteCarloPower(mt,param,1e4)
```

## Circadian study with random measurements

Same measurement budget as above but now measurements are collected at random time intervals
```{r}
mt_rand = runif(N)
evalExactPower(mt_rand,param)
```


## Worst case power
Given an experimental design, find the lowest power across all acrophases for signals of a fixed frequency and amplitude.
```{r}
evalWorstPower(mt,param)
```

As the frequency approaches the Nyquist rate, we should expect worst-case power of an equispaced design to drop to the type I error rate
```{r}
fNyq=4 # Nyquist rate of 8 measurement equispaced
scales = c(.5,.8,.9,.95,.99,.999)
for (scale in scales){
  param$freq=fNyq*scale
  wcp=evalWorstPower(mt,param)
  print(paste0('At ', 100*scale, '% of Nyquist freq, worst-case power = ',round(wcp,3)))
}
```


## Solve for sample size

Assume a study wants to use equispaced and detect signals of amplitude > 1 with at least 80% power
```{r}
thresh      = .8
N           = 4
freq        = 1
Amp         = 1
param       = list(freq=freq,Amp=Amp)
low_power   = T
while (low_power){
  mt        = c(1:N)/N - 1/N
  wcp       =  evalWorstPower(mt,param)
  low_power = evalWorstPower(mt,param)<thresh
  N         = N+1
}
print(paste0('Design with ', N,' measurements achieves ', round(wcp,4) , ' worst-case power.'))
```
